---
title: "week13ipPart2"
author: "Samuel Chege"
date: "7/19/2020"
output: html_document
---

1. Defining the Question

#a) Specifying the Data Analytic Question

Kira Plastinina is a Russian brand that is sold through a defunct chain of retail stores in Russia, Ukraine, Kazakhstan, Belarus, China, Philippines, and Armenia. The brand’s Sales and Marketing team would like to understand their customer’s behavior from data that they have collected over the past year. More specifically, they would like to learn the characteristics of customer groups.

#b) Defining the Metric for Success

* Perform clustering stating insights drawn from your analysis and visualizations.
* Upon implementation, provide comparisons between the approaches learned this week i.e. K-Means clustering vs Hierarchical clustering         highlighting the strengths and limitations of each approach in the context of your analysis.

#c) Understanding the context

understanding customer behavior is relevant to a business since it helps in determining sales distributions

#d) Recording the Experimental Design

Problem Definition
Data Sourcing
Check the Data
Perform Data Cleaning
Perform Exploratory Data Analysis  (Univariate, Bivariate & Multivariate)
Implement the Solution
Challenge the Solution
Follow up Questions

#e) Data Relevance

The dataset for this Independent project can be found here [http://bit.ly/EcommerceCustomersDataset

The dataset consists of 10 numerical and 8 categorical attributes.

 "Administrative"          
 "Administrative_Duration" 
 "Informational"           
 "Informational_Duration" 
 "ProductRelated"          
 "ProductRelated_Duration" 
 "BounceRates"             
 "ExitRates"              
 "PageValues"              
 "SpecialDay"              
 "Month"                   
 "OperatingSystems"       
"Browser"                 
"Region"                  
"TrafficType"             
"VisitorType"            
"Weekend"                 
"Revenue"

"Administrative", "Administrative Duration", "Informational", "Informational Duration", "Product Related" and "Product Related Duration" represents the number of different types of pages visited by the visitor in that session and total time spent in each of these page categories. The values of these features are derived from the URL information of the pages visited by the user and updated in real-time when a user takes an action, e.g. moving from one page to another. 
The "Bounce Rate", "Exit Rate" and "Page Value" features represent the metrics measured by "Google Analytics" for each page in the e-commerce site. 
The value of the "Bounce Rate" feature for a web page refers to the percentage of visitors who enter the site from that page and then leave ("bounce") without triggering any other requests to the analytics server during that session. 
The value of the "Exit Rate" feature for a specific web page is calculated as for all pageviews to the page, the percentage that was the last in the session.
The "Page Value" feature represents the average value for a web page that a user visited before completing an e-commerce transaction. 
The "Special Day" feature indicates the closeness of the site visiting time to a specific special day (e.g. Mother’s Day, Valentine's Day) in which the sessions are more likely to be finalized with the transaction. The value of this attribute is determined by considering the dynamics of e-commerce such as the duration between the order date and delivery date. For example, for Valentine’s day, this value takes a nonzero value between February 2 and February 12, zero before and after this date unless it is close to another special day, and its maximum value of 1 on February 8. 
The dataset also includes the operating system, browser, region, traffic type, visitor type as returning or new visitor, a Boolean value indicating whether the date of the visit is weekend, and month of the year.

2. Reading the Data

```{r}
library(corrplot)
library(ggplot2) #Plotting
library(dplyr)
library(tidyverse)
library(DataExplorer)
library(factoextra)
library(cluster)
```


```{r}
head(online_shoppers_intention)
```


```{r}
df <- data.frame(online_shoppers_intention)
head(df)
```

3. Checking the Data
```{r}
# Determining the no. of records in our dataset
dim(df)
```

```{r}
# Checking whether each column has an appropriate datatype
str(df)
```
```{r}
names(df)
```

5. Tidying the Dataset

# Checking for Outliers
```{r}
boxplot(df$Administrative)
```

```{r}
boxplot(df$Administrative_Duration)
```


```{r}
boxplot(df$Informational)
```


```{r}
boxplot(df$Informational_Duration)
```


```{r}
boxplot(df$ProductRelated)
```


```{r}
boxplot(df$ProductRelated_Duration)
```


```{r}
boxplot(df$BounceRates)
```


```{r}
boxplot(df$ExitRates)
```


```{r}
boxplot(df$PageValues)
```


```{r}
boxplot(df$SpecialDay)
```


```{r}
boxplot(df$OperatingSystems)
```


```{r}
boxplot(df$Browser)
```


```{r}
boxplot(df$Region)
```


```{r}
boxplot(df$TrafficType)
```
###we have outliers in our dataset but we won't deal with them now since the data maybe relevant

# Identifying the Missing Data

```{r}
colSums(is.na(df))
```
###we have some missing data in some columns

```{r}
#we fill the missing with their mean
df$Administrative[is.na(df$Administrative)] <- mean(df$Administrative, na.rm = TRUE)
df$Administrative_Duration[is.na(df$Administrative_Duration)] <- mean(df$Administrative_Duration, na.rm = TRUE)
df$Informational[is.na(df$Informational)] <- mean(df$Informational, na.rm = TRUE)
df$Informational_Duration[is.na(df$Informational_Duration)] <- mean(df$Informational_Duration, na.rm = TRUE)
df$ProductRelated[is.na(df$ProductRelated)] <- mean(df$ProductRelated, na.rm = TRUE)
df$ProductRelated_Duration[is.na(df$ProductRelated_Duration)] <- mean(df$ProductRelated_Duration, na.rm = TRUE)
df$BounceRates[is.na(df$BounceRates)] <- mean(df$BounceRates, na.rm = TRUE)
df$ExitRates[is.na(df$ExitRates)] <- mean(df$ExitRates, na.rm = TRUE)
```

```{r}
#confirming the missing data
colSums(is.na(df))
```

# Checking statistical summary of the dataset
```{r}
summary(df)
```

# checking for 
```{r}

numeric_data = df[, sapply(df, is.numeric)]

```

```{r}
corrplot(cor(numeric_data), method = 'shade')
```


6. Analysis

#univariate
```{r}
plot_density(df)
```


```{r}
plot_histogram(df,ncol = 4L)
```


```{r}

ggplot(data = df) +
  geom_bar(mapping = aes(x = VisitorType))
```


```{r}
ggplot(data = df) +
  geom_bar(mapping = aes(x = Weekend))
```


```{r}
ggplot(data = df) +
  geom_bar(mapping = aes(x = Revenue))
```
```{r}

plot(df$Administrative,df$Daily.Administrative_Duration, xlab='Administrative',ylab='Administrative_Duration')
```


```{r}
plot(df$Informational,df$Informational_Duration, xlab='Informational',ylab='Informational_Duration')
```


```{r}
plot(df$ProductRelated,df$ProductRelated_Duration, xlab='ProductRelated',ylab='ProductRelated_Duration')
```


```{r}
plot(df$BounceRates,df$ExitRates, xlab='BounceRates',ylab='ExitRates')
```

```{r}
plot(df$OperatingSystems,df$Browser, xlab='OperatingSystems',ylab='Browser')
```


```{r}
plot(df$Region,df$TrafficType, xlab='Region',ylab='TrafficType')
```

changing categorical values to numbers
```{r}

Month_enc = data.frame(model.matrix(~0+df$Month))
VisitorType_enc = data.frame(model.matrix(~0+df$VisitorType))
Weekend_enc = data.frame(model.matrix(~0+df$Weekend))
Revenue_enc = data.frame(model.matrix(~0+df$Revenue))

# Dropping non numerical columns 
drop_cols = c('Month', 'VisitorType','Weekend','Revenue')
df_customers = select(data.frame(cbind(df,Month_enc, VisitorType_enc,Weekend_enc,Revenue_enc)), -drop_cols)
```



MODELING
#K-MEANS CLUSTERING
```{r}
fviz_nbclust(df_customers, kmeans, method = "wss") +
    geom_vline(xintercept = 4, linetype = 2)+
  labs(subtitle = "Elbow method")
```


```{r}
kmeans_model = kmeans(df_customers, 4)

fviz_cluster(kmeans_model, df_customers)
```

#HIERACHICAL CLUSTERING
```{r}
df_H <- scale(df_customers)
```


```{r}
d <- dist(df_H, method = "euclidean")
```

```{r}
res.hc <- hclust(d, method = "ward.D2" )
```

```{r}
plot(res.hc, cex = 0.6, hang = -1)
```

